---
title: "Writing Functions"
author: "Alice Tivarovsky"
date: "10/24/2019"
output: html_document
---
**NOTE: Not connected to github**

# Slides

If you're using the same code more than once, write a function. Prevents mistakes from contnuous copy and paste. 

Functions consist of 
- arguments(inputs)
- body (code that does stuff)
- Return objects (what the function produces)
 
 Scoping: this is how R looks for variables. Usually we're in the global environment. With functions, we crat a local environment. 
 
 
# Example 

## Setup 
```{r setup, include=FALSE}

library(tidyverse)

```

## Get Started 

```{r}
x = rnorm (n = 30, mean = 4, sd = 2.3)
x_again = rnorm (n = 30, mean = 6, sd = .3)

z_transformation = mean(x)/sd(x)
(x_again - mean(x_again)/sd(x_again))
```

To avoid the above, write a function: 

```{r}
# here, x is what we're calling whatever we put in the function. 

z_score = function(x) {
  z = (x - mean(x))/sd(x)
  z
}
```

Now we try out the function. 

```{r}
z_score(x)

z_score(x = 3)
#z_score(x = "my name is alice")
z_score(x = c(TRUE, TRUE, FALSE, TRUE))

```

We can now define when the function breaks, i.e. conditional execution. 

```{r}
z_score = function(x) {
  
  if(!is.numeric(x)) {
    stop("argument should be numeric")}
  else if (length(x) == 1) {
    stop("Z scores cannot be computed for length 1 vectors")
  }
  z = (x - mean(x))/sd(x)
  z
}
```

Running again: 
```{r}
z_score(x)

#z_score(x = 3)
#z_score(x = "my name is alice")
#z_score(x = c(TRUE, TRUE, FALSE, TRUE))

```


## Multiple Outputs

Write a function that produces multiple outputs: 

```{r}

mean_and_sd = function(input_x) {
  
  if(!is.numeric(input_x)){
    stop("argument should be numeric")
    }
  else if (length(input_x) < 3) {
    stop("x should be larger than 3")
  }
  mean_x = mean(input_x)
  sd_x = sd(input_x)
}
```

Some stuff with lists. 

## Multiple Inputs

Go through this again. 

```{r}
sim_data = tibble(
  x = rnorm(30, mean = 1, sd = 1),
  y = 2 + 3 * x + rnorm(30, 0, 1)
)

ls_fit = lm(y ~ x, data = sim_data)
  
beta0_hat = coef(ls_fit)[1]
beta1_hat = coef(ls_fit)[2]
```


Write a function that simulates regression: 

```{r}

sim_regression = function(n, beta0 = 2, beta1 = 3) {
  
sim_data = tibble(
    x = rnorm(n, mean = 1, sd = 1),
    y = beta0 + beta1 * x + rnorm(n, 0, 1)
  )
  
ls_fit = lm(y ~ x, data = sim_data)
  
tibble(
    beta0_hat = coef(ls_fit)[1],
    beta1_hat = coef(ls_fit)[2]
  )
  }

sim_regression(3000)
```


# Argument names and scoping 

```{r}

```

