---
title: "Iteration and Listcols"
author: "Alice Tivarovsky"
date: "10/31/2019"
output: html_document
---


# Slides

###For loops 
Loops consist of an output object; a sequence to iterate over; the loop body; and (optionally) an input object. For loops are hard to work with but purr package makes for loops easier using maps. 

Basic structure: output = map(input, f)


###Lists
Lists hold everything. They give you a way to mix different data types (dbl vectors, string vectors, data frames, etc.). Dataframes are a special type of list - one that is formatted in a rectangle. 

You can even put lists inside of lists - these are called list columns. This makes data easier to work with (even though it's hard to understand in the beginning). 


#Examples 

Setup code: 
```{r setup, include=FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Lists
We need a brief digression about lists before we do anything.

In R, vectors are limited to a single data class – all elements are characters, or all numeric, or all logical. Trying to join the following vectors will result in _coersion_, as would creating vectors of mixed types.

```{r}
l = list(vec_numeric = 5:8,
         mat         = matrix(1:8, 2, 4),
         vec_logical = c(TRUE, FALSE),
         summary     = summary(rnorm(1000)))
l
```


 