---
title: "Simulations"
author: "Alice Tivarovsky"
date: "10/31/2019"
output: html_document
---


# Slides

Simulations are useful for statistical modeling. 




#Examples 

Setup code: 
```{r setup, include=FALSE}

library(tidyverse)

set.seed(1)

```

##Simulation: Simple Linear Regression for one n

```{r}
sim_regression = function(n, beta0 = 2, beta1 = 3) {
  
  sim_data = tibble(
    x = rnorm(n, mean = 1, sd = 1),
    y = beta0 + beta1 * x + rnorm(n, 0, 1)
  )
  
  ls_fit = lm(y ~ x, data = sim_data)
  
  tibble(
    beta0_hat = coef(ls_fit)[1],
    beta1_hat = coef(ls_fit)[2]
  )
}
```

Let’s run sim_regression() 100 times to see the effect of randomness in ϵ on estimates β̂0,β̂1

```{r}
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_regression(30)
}

sim_results = bind_rows(output)
```
 
Taking a look at the for loop we used to create these results, you might notice that there’s no input list – the sequence is used to keep track of the output but doesn’t affect the computation performed inside the for loop. In cases like these, the purrr::rerun function is very handy.


```{r}
sim_results = 
  rerun(100, sim_regression(30, 2, 3)) %>% 
  bind_rows()
```

Structurally, rerun is a lot like map – the first argument defines the amount of iteration and the second argument is the function to use in each iteration step. As with map, we’ve replaced a for loop with a segment of code that makes our purpose much more transparent but both approaches give the same results.

Let’s make some quick plots and compute some summaries for our simulation results.

```{r}
sim_results %>% 
  ggplot(aes(x = beta0_hat, y = beta1_hat)) + 
  geom_point()
```

